---
title: "Fast, Faster, ???"
subtitle: "Comparing Data Read/Write Speeds in R"
author: "David Zimmermann"
institute: "R User Group Cologne"
date: "2020-03-05 @ Cologne Intelligence"
output:
  xaringan::moon_reader:
    lib_dir: libs
    css: xaringan-themer.css
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---
class: center, middle, inverse

```{r xaringan-themer, include = FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(xaringanthemer)
suppressPackageStartupMessages(library(tidyverse))
library(kableExtra)
library(fontawesome)
library(formattable)

duo(
  primary_color = "#1F4257", 
  secondary_color = "#F97B64",
  extra_css = list(
    ".small" = list("font-size" = "80%"),
    ".smaller" = list("font-size" = "67%"),
    ".smallish" = list("font-size" = "60%"),
    ".smallest" = list("font-size" = "50%"),
    ".tiny" = list("font-size" = "20%"),
    ".scroll-output" = list("height" = "55em", "overflow-y" = "scroll"),
    ".remark-slide table" = list("border-bottom" = "none", "border-top" = "none", "width" = "100%"),
    ".move-upwards" = list("margin-top" = "-50px")
  ),
  table_border_color = "gray",
  table_row_border_color = "lightgray",
  table_row_even_background_color = NA,
  header_font_google = google_font("Josefin Sans"),
  text_font_google   = google_font("Montserrat", "300", "300i"),
  code_font_google   = google_font("Droid Mono")
)

options(htmltools.dir.version = FALSE)

# which datasets do we compare?
data_select <- c("tiny", "wide", "large", "nycflights", "tickets2018")

f_check <- function(x) map_chr(x, ~ifelse(is.na(.x) | .x != TRUE, "", as.character(fa("check", fill = "white"))))

# wraps something in code tags
as_code <- function(x) paste0("<code>", x, "</code>") 

# maps a vector of values to a map of fontawesome signs
fa_map <- function(x, map) {
  xx <- as.character(x)
  r <- map[xx]
  as.character(map_chr(r, ~ifelse(is.na(.x), "", fontawesome::fa(.x, fill = "white"))))
}

# override fa to accomodate vectors of x
fa <- function(x, height = NULL, fill = NULL) purrr::map_chr(x, ~fontawesome::fa(.x, height = height, fill = fill))

# formattable::color_tile but with the option to specify a function that is applied afterwards
color_tile_fix <- function(x, 
                           bg_colors = c("#e8554e", "#ffd265", "#2aa876"), 
                           font_colors = "black", 
                           font_size = 1, # in em
                           log_fmt = FALSE,
                           f = NA) {
  txt <- "<span style=\"display: block; width:100%;height:100%; font-size: {font_size}em; padding: 0 4px; border-radius: 4px; background-color: {bg_col}; color: {ft_col};\">{val}</span>"
  
  if (log_fmt) x <- log(x)
  
  bg_cols <- color_gradient(x, bg_colors)
  ft_cols <- color_gradient(x, font_colors)
  
  if (is.function(f)) {
    if (log_fmt) {
      x <- f(exp(x))
    } else {
      x <- f(x)
    }
  }
  
  purrr::pmap_chr(list(x, bg_cols, ft_cols), 
                  function(val, bg_col, ft_col) glue::glue(txt))
}

red_to_green <- c("#e8554e", "#ffd265", "#2aa876")
color_gradient <- function(x, colors = c("#e8554e", "#ffd265", "#2aa876"), colsteps = 100)
  colorRampPalette(colors)(colsteps)[findInterval(x, seq(min(x, na.rm = TRUE), max(x, na.rm = TRUE), length.out = colsteps))]

pretty_bytes_sec <- function(x) paste0(prettyunits::pretty_bytes(x, style="6"), "/s")

fix_names <- function(d) {
  n <- stringr::str_replace(names(d), "^(speed|RAM|filesize|read|write)_", "")
  n[n == "tickets2018"] <- "tickets"
  n[n == "nycflights"] <- "flights"
  names(d) <- n
  d
}
files <- list.files("results", ".qs$", full.names = TRUE)
df_raw <- map_dfr(files, qs::qread) %>%
  rename(Library = library, Fileformat = filetype) %>% 
  filter(dataset %in% data_select)

od <- c("tiny" = 1, "small" = 2, "wide" = 3, "medium" = 4, "large" = 5, "nycflights" = 6, "tickets2018" = 7)
obj_sizes <- df_raw %>% 
  group_by(dataset) %>% 
  summarise(object_size = mean(object_size),
            file_bytes = prettyunits::pretty_bytes(object_size, style = "6")) %>% 
  mutate(idx = od[dataset]) %>% 
  arrange(idx)

order <- tibble(
  Library = c("base", "readr", "data.table", "vroom", "base", "readr", "fst",
              "qs", "feather", "arrow", "arrow", "arrow"), 
Fileformat = c(".csv", ".csv", ".csv", ".csv", ".rds", ".rds", ".fst", 
               ".qs", ".feather", ".feather", ".arrow", ".parquet"), 
  id = 1:12)

df <- df_raw %>% 
  filter(is.na(compression)) %>% 
  group_by(Library, Fileformat, dataset, operation) %>% 
  summarise(speed = mean(time), RAM = mean(mem_alloc), filesize = mean(filesize)) %>%
  ungroup() %>% 
  mutate(Library = ifelse(Library == "datatable", "data.table", Library),
         Fileformat = paste0(".", Fileformat),
         speed = ifelse(Library == "vroom" & operation == "read", NA, speed),
         RAM = ifelse(Library == "vroom" & operation == "read", NA, RAM)) %>% 
  left_join(order, by = c("Library", "Fileformat")) %>% 
  arrange(id) %>% select(-id)
```


# How can we write and read data in R efficiently?

--

### Or

## WHAT? ... you still use read.csv?

---

# Soooo many options...

.pull-left[
## Filetypes
- `.csv`
- `.rds`
- `.fst`
- `.qs`
- `.feather`
- `.arrow`
- `.parquet`
]
--
.pull-right[
## R packages
- `base`
- `readr`
- `data.table`
- `vroom`
- `fst`
- `qs`
- `feather`
- `arrow`
]

--

**+** different compression levels

---

# Multitude of wonderful packages

```{r, echo=F}
libs <- tribble(~Library, ~format,
             "base", "csv",
             "base", "rds",
             "readr", "csv",
             "readr", "rds",
             "data.table", "csv",
             "vroom", "csv",
             "fst", "fst",
             "qs", "qs",
             "feather", "feather",
             "arrow", "feather",
             "arrow", "arrow",
             "arrow", "parquet") %>% 
  mutate(format = paste0("<code>.", format, "</code>"), val = TRUE) %>% 
  pivot_wider(names_from = format, values_from = val, values_fill = list(val = FALSE)) %>% 
  mutate(Library = as_code(Library)) %>%
  mutate_at(vars(-one_of("Library")), f_check) %>% 
  select(Library, everything())

libs %>%
  kable(escape = FALSE, align = "lcccccccc") %>%
  kable_styling() %>% 
  add_header_above(c(" ", "Filetypes" = 7))
```


---
class: center, middle, inverse
# And the winner is?

--

# qs

--

# ... but, its more complicated

---

# The many factors to consider

.pull-left[

### Factors to consider
- **RAM** usage
- Read/Write **speed**
- **Filesize**

depends on ... 
]

--

.pull-right[

### Type of data
- **size** (KBs, MBs, GBs)
- **shape** (long, wide)
- **type** (string, int, numeric, lists in dataframes)

`+`

### Additional factors
- Python **compatiblity**/file formats
- **Dependency** weight
- **Partial read** (row / column wise)
- Support of **nested data-frames**

]

---
class: center, middle, inverse

# Just give me one option to use!

---
# The Real Winner is

I want to ...

<table style="width:100%;">
  <tr>
    <th style="text-align: left;"><li> have the <strong>fastest</strong> read/writes: </li></th>
    <th style="text-align: left;"><code> fst </code></th>
  </tr>
  <tr>
    <th style="text-align: left;"><li> have the smallest <strong>RAM usage</strong>: </li></th>
    <th style="text-align: left;"><code> qs </code></th>
  </tr>
  <tr>
    <th style="text-align: left;"><li> have the smallest <strong>filesize on disk</strong>: </li></th>
    <th style="text-align: left;"><code> qs </code> or <code>base->rds</code> (slow...)</th>
  </tr>
  <tr>
    <th style="text-align: left;"><li> <strong>communicate</strong> with everything: </li></th>
    <th style="text-align: left;"><code> data.table </code></th>
  </tr>
  <tr>
    <th style="text-align: left;"><li> have light <strong>dependencies</strong>: </li></th>
    <th style="text-align: left;"><code>base</code> or <code>data.table</code></th>
  </tr>
  <tr>
    <th style="text-align: left;"><li> also use it in <strong>python</strong>: </li></th>
    <th style="text-align: left;"><code> arrow->feather </code></th>
  </tr>
  <tr>
    <th style="text-align: left;"><li>  use <strong>nested dataframes</strong>: </li></th>
    <th style="text-align: left;"><code>qs</code></th>
  </tr>
</table>
    
```{r dependencies, include=FALSE, cache=TRUE}
pkgs <- c("base", "readr", "data.table", "vroom", "qs", "fst", "feather", "arrow")
deps <- tools::package_dependencies(pkgs, recursive = TRUE)
deps
```

<br>

Source code can be found here: [https://github.com/DavZim/io-benchmarks](https://github.com/DavZim/io-benchmarks)

---
### Summary

.smaller[
```{r, echo=FALSE}
mults <- df_raw %>% 
  filter(is.na(compression)) %>% 
   select(Library:compression, 
         speed = time, RAM = mem_alloc, filesize) %>% 
  group_by(operation, dataset) %>%
  mutate(RAM = RAM + 1) %>% # avoid div/0
  mutate_at(vars(speed, RAM, filesize), list(mult = ~.x/min(.x))) %>% 
  group_by(Library, Fileformat) %>% 
  summarise_at(vars(ends_with("_mult")), mean, na.rm = TRUE) %>% 
  ungroup() %>% 
  select(Library, Fileformat, Speed = speed_mult, RAM = RAM_mult, Filesize = filesize_mult) %>% 
  mutate(Fileformat = paste0(".", Fileformat),
         Library = ifelse(Library == "datatable", "data.table", Library)) %>% 
  mutate_at(vars(Speed, RAM, Filesize), list(~ifelse(Library == "vroom", NA, .x)))

# 0 -> NO, 1 -> YES, 2 -> Best in Class, 9 -> unclear/future (maybe), 3 -> row only random access
# Dependencies: 0: best in class, 1: some=good, 2: many=bad
# Speed, RAM, size: 0: best in class ("medal"), 1: good, 2: average, 3: bad
spec_res <- tribble(
  ~Library, ~Fileformat, ~Multicore, ~`Nested DFs`, ~Python, ~`Random Access`, ~Dependencies,
  "base",        ".csv",      FALSE,         FALSE,       1,                0,             0,
  "readr",       ".csv",      FALSE,         FALSE,       1,                0,             2,
  "data.table",  ".csv",       TRUE,         FALSE,       1,                3,             0,
  "vroom",       ".csv",       TRUE,         FALSE,       1,                3,             2,
  "base",        ".rds",      FALSE,          TRUE,       0,                0,             0,
  "readr",       ".rds",      FALSE,          TRUE,       0,                0,             2,
  "fst",         ".fst",       TRUE,         FALSE,       9,                2,             1,
  "qs",           ".qs",       TRUE,          TRUE,       0,                0,             1,
  "feather", ".feather",      FALSE,         FALSE,       1,                0,             1,
  "arrow",   ".feather",      FALSE,         FALSE,       1,                3,             2,
  "arrow",     ".arrow",      FALSE,         FALSE,       1,                0,             2,
  "arrow",   ".parquet",      FALSE,         FALSE,       1,                3,             2
  ) %>% 
  left_join(mults, by = c("Library", "Fileformat"))

# maps that say how values are decoded to fontawesome glyphs
python_map <- c("0" = NA, "1" = "check", "9" = "question")
ra_map <- c("0" = NA, "2" = "crown", "3" = "angle-down")
dep_map <- c("0" = "crown", "1" = "thumbs-up", "2" = "hand-point-right")
fac_map <- c("0" = "crown", "1" = "thumbs-up", "2" = "hand-point-right", "3" = "thumbs-down")
rank_map <- function(x) x %>% percent_rank() %>% {ifelse(. < 0.01, "crown", 
                                                        ifelse(. < 0.3, "thumbs-up",
                                                               ifelse(. < 0.7, "hand-point-right", "thumbs-down")))} %>% fa()

spec_res %>% 
  mutate(
    Library = as_code(Library),
    Fileformat = as_code(Fileformat),
    Multicore = f_check(Multicore),
    `Nested DFs` = f_check(`Nested DFs`),
    Python = fa_map(Python, python_map),
    `Random Access` = fa_map(`Random Access`, ra_map),
    Dependencies = fa_map(Dependencies, dep_map),
    Speed = color_tile_fix(Speed, rev(red_to_green), f = rank_map),
    RAM = color_tile_fix(RAM, rev(red_to_green), log_fmt = TRUE, f = rank_map),
    Filesize = color_tile_fix(Filesize, rev(red_to_green), f = rank_map)
    ) %>% 
  kable(escape = FALSE, align = "lrccccclll") %>%
  kable_styling() %>% 
  pack_rows("Plain Text", 1, 4) %>%
  pack_rows("R specific", 5, 8) %>%
  pack_rows("Arrow-based", 9, 12) %>%
  add_header_above(c(" " = 2, "Capabilities" = 5, "Factors" = 3))
```
]

.smaller[Note that `vroom` is not properly benchmarked, due to lazy loading(?!)]

---
# Special Mentions & Remarks

### Special Mentions
.smaller[
- `data.table` for its use of pipes: (search all `.csv.gz` archives for the line `rug-meetup` and load only those lines)

```r
fread(cmd = "cat *.csv.gz | gunzip | grep -e rug-meetup")
``` 
]

--

.smaller[
- `vroom` for its excellent functionality and performance with strings
]

--

.smaller[
- `qs` for its excellent compression, i.e., `preset="archive"` reduces the `tickets` dataset from 3.4 GB in `R` to 278 MB (!) on disk in ~3 mins or in 12 secs to 580 MB with `preset="balanced"`

```r
qsave(tickets, "tickets.qs", preset = "archive")
```
]

--

### Remarks

.smaller[
- `vroom` loads the data only lazily, resulting in inaccurate measurements (i.e., 2 MB when GBs where used on the machine), it is a lot better than these benchmarks suggest.
]

--

.smaller[
- `qs` allows multicore, but default is to only use 1, for using all available cores, use

```r
qs::qread("myfile.qs", nthrads = parallel::detectCores())
```

]


---
class: center, middle, inverse

# Thats it for today

--

# ... OK, ok, BUT I still want the details...

### see following slides
---

### Average Speed
.smallish[
```{r, echo=FALSE, warning=FALSE}
ll <- list(speed    = df %>% select(Library:operation, val = speed),
           RAM      = df %>% select(Library:operation, val = RAM),
           filesize = df %>% select(Library:operation, val = filesize)) %>%
  # pivot values to wider
  map(
  ~pivot_wider(.x, Library:operation, names_from = dataset, values_from = val) %>%
    select(Library, Fileformat, operation, one_of(data_select))
  ) %>% 
  # split read/write
  map(function(data) {
    split <- split(data, data$operation)
    
    split[["read"]] <- split[["read"]] %>% 
      rename_at(vars(one_of(data_select)), ~paste0("read_", .x)) %>%
      select(-operation)
    
    split[["write"]] <- split[["write"]] %>% 
      rename_at(vars(one_of(data_select)), ~paste0("write_", .x)) %>% 
      select(-operation)
    
    Reduce(function(x, y) left_join(x, y, by = c("Library", "Fileformat")), split)
  })

ll$speed %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = prettyunits::pretty_sec)) %>%
  mutate(Library = as_code(Library)) %>%
  fix_names() %>%
  set_names(c("Library", "Fileformat", rep(obj_sizes$file_bytes, 2))) %>% 
  kable(escape = FALSE, align = "lrllllllllllllllllllllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("Plain Text", 1, 4, bold = TRUE) %>%
  pack_rows("R specific", 5, 8, bold = TRUE) %>%
  pack_rows("Arrow-based", 9, 12, bold = TRUE) %>%
  add_header_above(c(" " = 2,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  add_header_above(c(" " = 2, 
                     "Read" = length(data_select),
                     "Write" = length(data_select))) %>%
  column_spec(c(1, 2), bold = TRUE)
```
]

.smaller[Note that all comparisons are performed without compression. Also, note that `vroom` benchmarks are not accurate for read-operations]

---
### Average RAM Usage
.smallish[
```{r, echo=FALSE}
ll$RAM %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x + 1, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = function(x) prettyunits::pretty_bytes(x, style="6"))) %>%
  mutate(Library = as_code(Library)) %>%
  fix_names() %>%
  set_names(c("Library", "Fileformat", rep(obj_sizes$file_bytes, 2))) %>% 
  kable(escape = FALSE, align = "lrllllllllllllllllllllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("Plain Text", 1, 4, bold = TRUE) %>%
  pack_rows("R specific", 5, 8, bold = TRUE) %>%
  pack_rows("Arrow-based", 9, 12, bold = TRUE) %>%
  add_header_above(c(" " = 2,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  add_header_above(c(" " = 2, 
                     "Read" = length(data_select),
                     "Write" = length(data_select))) %>%
  column_spec(c(1, 2), bold = TRUE)
```
]

.smaller[Note that all comparisons are performed without compression. Also, note that `vroom` benchmarks are not accurate for read-operations]
---
### Average Filesize
.smallish[
```{r, echo=FALSE}
ll$filesize %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x + 1, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = function(x) prettyunits::pretty_bytes(x - 1, style="6"))) %>%
  mutate(Library = as_code(Library)) %>%
  select(Library, Fileformat, starts_with("write")) %>% 
  fix_names() %>%
  set_names(c("Library", "Fileformat", obj_sizes$file_bytes)) %>% 
  kable(escape = FALSE, align = "lrllllllllllllllllllllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("Plain Text", 1, 4, bold = TRUE) %>%
  pack_rows("R specific", 5, 8, bold = TRUE) %>%
  pack_rows("Arrow-based", 9, 12, bold = TRUE) %>%
  add_header_above(c(" " = 2,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  column_spec(c(1, 2), bold = TRUE)
```
]
.smaller[Note that all comparisons are performed without compression]
---
class: center, middle, inverse

# What about compressions?

---

### Average Speed Compression
.move-upwards[.smallish[
```{r, echo=FALSE}
df <- df_raw %>% 
  filter(!is.na(compression)) %>% 
  group_by(Library, Fileformat, dataset, operation, compression) %>% 
  summarise(speed = mean(time), RAM = mean(mem_alloc), filesize = mean(filesize)) %>%
  ungroup() %>% 
  mutate(Library = ifelse(Library == "datatable", "data.table", Library),
         Fileformat = paste0(".", Fileformat)) %>% 
  left_join(order, by = c("Library", "Fileformat")) %>% 
  arrange(id) %>% select(-id) %>% 
  rename(Compression = compression)

ll <- list(speed    = df %>% select(Library:Compression, val = speed),
           RAM      = df %>% select(Library:Compression, val = RAM),
           filesize = df %>% select(Library:Compression, val = filesize)) %>%
  # pivot values to wider
  map(
  ~pivot_wider(.x, Library:Compression, names_from = dataset, values_from = val) %>%
    select(Library, Fileformat, operation, Compression, one_of(data_select))
  ) %>% 
  # split read/write
  map(function(data) {
    split <- split(data, data$operation)
    
    split[["read"]] <- split[["read"]] %>% 
      rename_at(vars(one_of(data_select)), ~paste0("read_", .x)) %>%
      select(-operation)
    
    split[["write"]] <- split[["write"]] %>% 
      rename_at(vars(one_of(data_select)), ~paste0("write_", .x)) %>% 
      select(-operation)
    
    Reduce(function(x, y) left_join(x, y, by = c("Library", "Fileformat", "Compression")), split)
  })

ll$speed %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = prettyunits::pretty_sec)) %>%
  mutate(Library = as_code(Library)) %>%
  fix_names() %>%
  set_names(c("Library", "Fileformat", "Compression", rep(obj_sizes$file_bytes, 2))) %>% 
  kable(escape = FALSE, align = "lrrllllllllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("<code>base</code>", 1, 3, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>readr</code>", 4, 7, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>fst</code>", 8, 9, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>qs</code>", 10, 12, bold = TRUE, escape = FALSE) %>%
  add_header_above(c(" " = 3,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  add_header_above(c(" " = 3, 
                     "Read" = length(data_select),
                     "Write" = length(data_select))) %>%
  column_spec(c(1, 2, 3), bold = TRUE)
```
]]

---
### Average RAM Usage Compression
.move-upwards[.smallish[
```{r, echo=F}
ll$RAM %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x + 1, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = function(x) prettyunits::pretty_bytes(x - 1, style = "6"))) %>%
  mutate(Library = as_code(Library)) %>%
  fix_names() %>%
  set_names(c("Library", "Fileformat", "Compression", rep(obj_sizes$file_bytes, 2))) %>% 
  kable(escape = FALSE, align = "lrrllllllllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("<code>base</code>", 1, 3, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>readr</code>", 4, 7, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>fst</code>", 8, 9, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>qs</code>", 10, 12, bold = TRUE, escape = FALSE) %>%
  add_header_above(c(" " = 3,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  add_header_above(c(" " = 3, 
                     "Read" = length(data_select),
                     "Write" = length(data_select))) %>%
  column_spec(c(1, 2, 3), bold = TRUE)
```
]]

---
###  Average Filesize Compression
.smallish[
```{r, echo=F}
ll$filesize %>%
  mutate_if(is.numeric,
            ~color_tile_fix(.x + 1, rev(red_to_green), 
                            font_colors = "black", font_size = 0.75, 
                            log_fmt = TRUE,
                            f = function(x) prettyunits::pretty_bytes(x - 1, style="6"))) %>%
  mutate(Library = as_code(Library)) %>%
  select(Library, Fileformat, Compression, starts_with("write")) %>% 
  fix_names() %>%
  set_names(c("Library", "Fileformat", "Compression", obj_sizes$file_bytes)) %>% 
  kable(escape = FALSE, align = "lrrlllll") %>%
  kable_styling("hover", full_width = TRUE) %>%
  pack_rows("<code>base</code>", 1, 3, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>readr</code>", 4, 7, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>fst</code>", 8, 9, bold = TRUE, escape = FALSE) %>%
  pack_rows("<code>qs</code>", 10, 12, bold = TRUE, escape = FALSE) %>%
  add_header_above(c(" " = 3,
                     "tiny" = 1, "wide" = 1, "large" = 1, "flights" = 1, "tickets" = 1)) %>%
  column_spec(c(1, 2, 3), bold = TRUE)
```
]

---
class: center, middle, inverse

# Code Examples

---
# Code Examples

.pull-left[.smaller[
### Plain text file formats
Package `base`: [`write.csv()`](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/write.table) & [`read.csv()`](https://www.rdocumentation.org/packages/utils/versions/3.6.2/topics/read.table)
```{r, eval=FALSE}
write.csv(df, "dataset.csv")
df2 <- read.csv("dataset.csv")
```
Package [`readr`](https://readr.tidyverse.org/): [`write_csv()`](https://www.rdocumentation.org/packages/readr/versions/1.3.1/topics/write_delim) & [`read_csv()`](https://www.rdocumentation.org/packages/readr/versions/1.3.1/topics/read_delim)
```{r, eval=FALSE}
library(readr)
write_csv(df, "dataset.csv")
df2 <- read_csv("dataset.csv")
```
Package [`data.table`](https://github.com/Rdatatable/data.table): [`fwrite()`](https://www.rdocumentation.org/packages/data.table/versions/1.12.8/topics/fwrite) & [`fread()`](https://www.rdocumentation.org/packages/data.table/versions/1.12.8/topics/fread)
```{r, eval=FALSE}
library(data.table)
fwrite(df, "dataset.csv")
df2 <- fread("dataset.csv")
```
Package [`vroom`](https://github.com/r-lib/vroom): [`vroom_write()`](https://www.rdocumentation.org/packages/vroom/versions/1.0.2/topics/vroom_write) & [`vroom()`](https://www.rdocumentation.org/packages/vroom/versions/1.0.2/topics/vroom)
```{r, eval=FALSE}
library(vroom)
vroom_write(df, "dataset.csv")
df2 <- vroom("dataset.csv")
```
]]

.pull-right[.smaller[
### `R` specific file formats
Package `base`: [`saveRDS()` & `readRDS()`](https://www.rdocumentation.org/packages/base/versions/3.6.2/topics/readRDS) 
```{r, eval=FALSE}
saveRDS(df, "dataset.rds", compress = "none")
df2 <- readRDS("dataset.rds")
```
Package [`readr`](https://readr.tidyverse.org/): [`write_rds()` & `read_rds()`](https://www.rdocumentation.org/packages/readr/versions/1.3.1/topics/read_rds)
```{r, eval=FALSE}
library(readr)
write_rds(df, "dataset.rds", compress = "none")
df2 <- read_rds("dataset.rds")
```
Package [`fst`](https://www.fstpackage.org/): [`fst_write()` & `fst_read()`](https://www.rdocumentation.org/packages/fst/versions/0.9.0/topics/write_fst)
```{r, eval=FALSE}
library(fst)
write_fst(df, "dataset.fst", compression = 50)
df2 <- read_fst("dataset.fst")
```
Package [`qs`](https://github.com/traversc/qs): [`qsave()`](https://www.rdocumentation.org/packages/qs/versions/0.21.1/topics/qsave) & [`qread()`](https://www.rdocumentation.org/packages/qs/versions/0.21.1/topics/qread)
```{r, eval=FALSE}
library(qs)
qsave(df, "dataset.qs", preset = "high")
df2 <- qread("dataset.qs")
```
]]

---
# Code Examples cont'd

### `arrow` based file formats
.pull-left[.smaller[
Package [`feather`](https://github.com/wesm/feather): [`write_feather()` & `read_feather()`](https://www.rdocumentation.org/packages/feather/versions/0.3.5/topics/read_feather)
```{r, eval=FALSE}
library(feather)
write_feather(df, "dataset.feather")
df2 <- read_feather("dataset.feather")
```
Package [`arrow`](https://github.com/apache/arrow/tree/master/r): [`write_feather()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/write_feather) & [`read_feather()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/read_feather)
```{r, eval=FALSE}
library(arrow)
# install_arrow()
write_feather(df, "dataset.feather")
df2 <- read_feather("dataset.feather")
```
]]
.pull-right[.smaller[
Package [`arrow`](https://github.com/apache/arrow/tree/master/r): [`write_arrow()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/write_arrow) & [`read_arrow()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/read_table)
```{r, eval=FALSE}
library(arrow)
# install_arrow()
write_arrow(df, "dataset.arrow")
df2 <- read_arrow("dataset.arrow")
```
Package [`arrow`](https://github.com/apache/arrow/tree/master/r): [`write_parquet()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/write_parquet) & [`read_parquet()`](https://www.rdocumentation.org/packages/arrow/versions/0.16.0.2/topics/read_parquet)
```{r, eval=FALSE}
library(arrow)
# install_arrow()
write_parquet(df, "dataset.parquet")
df2 < read_parquet("dataset.parquet")
```
]]

---
class: center, middle, inverse
# System Information

---

.pull-left[
## Hardware
- Dell XPS 15 Laptop
- CPU: Intel Core i7 (8th gen)
- Model: 9750H
- Clock speed: 2.60GHz 
- Threads: 6 cores, 12 threads
- RAM: 32 GB
- SSD with ~900 MB/sec read speed
]
.pull-right[
## Software Platform
.small[
```{r}
sessioninfo::platform_info()
```
]]

---
## Benchmark package versions
.smallest[
```{r}
pkgs <-  c("readr", "data.table", "vroom", "fst", "qs", "feather", "arrow")
sessioninfo::package_info(pkgs)
```
]